Sequential algorithms to split and merge ultra-high resolution 3D images

### 2017-07-30 paper.tex

\begin{algorithm}[h]
  \caption{Buffered merging of blocks with multiple reads}
  \label{algo:multiple-reads}
  \begin{algorithmic}[1]
  \STATE sorted\_split\_name\_list = sort blocks by increasing (z,y,x) values
  \STATE start\_index = 0, end\_index=(m-1)
  \WHILE{end\_index \textless sizeof(reconstructed\_image)}
    \STATE initialize data\_buffer
    \FOR{split\_name in sorted\_split\_name\_list}
      \IF{indexes of split\_name is in next\_write\_range}
        \STATE read split
        \STATE extract all indexes in next write range
        \STATE write the corresponding data to the data\_buffer
      \ENDIF
    \ENDFOR
    \STATE write data\_buffer in reconstructed\_image
    \STATE clear data\_buffer
    \STATE start\_index = end\_index + 1
    \STATE end\_index += mem
  \ENDWHILE

@@

@@

Experiment scripts are available at xxx.

>>>>

\begin{algorithm}[h]
  \caption{Buffered merging of blocks with multiple reads}
  \label{algo:multiple-reads}
  \begin{algorithmic}[1]
  \STATE sorted\_split\_list = sort splits based on its postion in increasing (z,y,x) values
  \STATE start\_index = 0, end\_index=(m-1)
  \STATE write\_range = (start\_index, end\_index)
  \WHILE{end\_index \textless sizeof(reconstructed\_image)}
    \STATE initialize data\_buffer
    \FOR{split in sorted\_split\_list}
      \IF{split is in write\_range}
        \STATE read split
        \STATE extract the rows in the write\_range
        \STATE write rows to the data\_buffer
      \ENDIF
    \ENDFOR
    \STATE write data\_buffer continuously to reconstructed\_image
    \STATE clear data\_buffer
    \STATE start\_index = end\_index + 1
    \STATE end\_index += mem
  \ENDWHILE

@@

For faster reading all the splits, before the merging process start,
we read all the split’s headers to get their position information and
calculate each row’s index in the reconstructed image. We only read the
split when there is a row in that writing range (available memory),
otherwise, we skip it.

Initially, we use a numpy array as our data buffer, after the split is read,
we assign the rows (a numpy array of unsigned integers) into this
buffer to the right position. However, after some tests, we found that
there is some overhead. Each time we assign the data to the corresponding
position of the buffer, the system will copy these integers first.
This is neither time-efficient nor space-efficient.

We then think of a better solution which we use the dictionary in python
as our data buffer. We use the row’s index as a key, the data (in bytes)
as the value. As for the multiple reads, the data buffer only contains
contiguous bytes, we can just seek once (the first key of the sorted keys
in dictionary), then write all the bytes continuously to the file.
There is no copy in the memory, which proves a bit faster than using
a numpy array as the buffer.

@@

Experiment scripts are available at xxx.
As our experiments results are heavily rely on IO performance of the disk,
we drop the system cache to get more precise result before each experiment.
